---
title: "Sesión 6"
output: html_document
date: "2022-08-13"
---
Grouping and summarizing 

Sumarizing the meidian life expectancy

```{r}
library(gapminder)
library(dplyr)
gapminder %>%
    summarize(medianLifeExp = median(lifeExp))
#Filter for 1957 then summarize the median life expectancy
gapminder %>%
    filter(year ==1957) %>%
    summarize(medianLifeExp = median(lifeExp))
gapminder %>%
    filter(year ==1957) %>%
    summarize(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))
```

Summarizing by year

```{r}
gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp), maxGdpPercap=max(lifeExp))
```

Summarizing by continent
```{r}
gapminder %>%
  filter(year == 1957) %>%
  group_by(continent) %>%
  summarize(medianLifeExp=median(lifeExp), maxGdpPercap=max(gdpPercap)) %>%
  arrange(desc(maxGdpPercap))
```

Summarizing by continent and year

```{r}
gapminder %>%
  group_by(continent,year) %>%
  summarize(medianLifeExp=median(lifeExp), maxGdpPercap=max(gdpPercap))
```

Visualizing median life expectancy over time
```{r}
by_year <- gapminder %>%
  group_by(year) %>%
  summarize(medianGdpPercap= median(gdpPercap))
by_year
ggplot(by_year, aes(x=year, y =medianGdpPercap))+
  geom_line()
  
```

```{r}
by_year_continent <- gapminder %>%
  group_by(year, continent) %>%
  summarize(medianGdpPercap=median(gdpPercap))
by_year_continent
ggplot(by_year_continent, aes(x=year,y=medianGdpPercap, color=continent))+
  geom_line()
```

Bar Plot

Visualizing median GDP per capita by continent

```{r}
by_continent <- gapminder %>%
  filter(year==1952) %>%
  group_by(continent) %>%
  summarize(medianGdpPercap = median(gdpPercap))
  by_continent
  
  ggplot(by_continent, aes(x=continent,y=medianGdpPercap))+
    geom_col()
```
Visualizing GDP per capita by country in Oceania

```{r}
oceania_1952 <- gapminder %>%
  filter(continent=="Oceania", year==1952)
  
oceania_1952  
ggplot(oceania_1952, aes(x=country, y=gdpPercap))+
  geom_col()
```

Histogram

Visualizing population

```{r}
library(gapminder)
library(ggplot2)
library(dplyr)
gapminder
gapminder_1952 <- gapminder %>%
  filter(year==1952) %>%
  mutate(pop_by_mil = pop/1000000)
gapminder_1952
ggplot(gapminder_1952, aes(x=pop_by_mil))+
  geom_histogram(bins=50)
```

Visualizing population with x-axis on a log scale

```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)
ggplot(gapminder_1952, aes(x=pop)) +
  geom_histogram(bins=50) +
  scale_x_log10()
```


Boxplots

Comparing GDP per capita across continents

```{r}
gapminder_1952 <- gapminder %>%
  filter(year==1952)
ggplot(gapminder_1952, aes(x=continent, y=gdpPercap))+
  geom_boxplot()+
  scale_y_log10()+
  ggtitle("Comparing GDP per capita")
```
Reading csv

```{r}
library(tidyverse)
counties <- read.csv('counties.csv')
glimpse(counties)
counties
```

select, filter, arrange

```{r}
counties %>%
  select(state, county, population, men, women)
counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)
counties_selected %>%
  arrange(desc(public_work))
counties_selected %>%
  select(state, county, population)
counties_selected %>%
  filter(population>1000000)
counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)
counties_selected %>%
  filter(state="Texs", population>1000) %>%
  arrange(desc(private_work))
```


Select, mutate, filter and arrange
```{r}
counties %>%
  select(state, county, population, men, women) %>%
  mutate(proportion_men = men/population) %>%
  filter(popultion>10000) %>%
  arrange(desc(proportion_men))
```


the verb count
```{r}
select_counties <- counties %>%
    select(state, region, county, population, citizens) 
counties %>%
  count(region, sort = TRUE) %>%
  arrange(desc(n))
```
Counting citizens by state

```{r}
counties_selected <-  counties %>%
  select(county, region, state, population, citizens)
counties_selected %>%
  count(state,wt=citizens) %>% #El conteo lo hace por medio de citizens(wt) no por state, por state solo los agrupa
  arrange(desc(n))
```

Mutating and counting

```{r}
counties_selected <-  counties %>%
  select(county, region, state, population, walk)
counties_selected %>%
  mutate(population_walk=population*walk/100) %>%
  count(state, wt=population_walk, sort=TRUE)
```

Summarizin

```{r}
counties_selected <- counties %>%
  select(county, population, income, unemployment)
counties_selected %>%
  summarize(min_population=min(population), max_unemploment=max(unemployment), average_income=mean(income))
```

Summarizing by state

```{r}
counties_selected <- counties %>%
  select(state,county, population, land_area)
counties_selected %>%
  group_by(state) %>%
  summarize(total_area=sum(land_area),total_population=sum(population)) %>%
  mutate(density=total_population/total_area) %>%
  arrange(desc(density))
counties_selected <- counties %>%
  select(region,state,county, population)
counties_selected %>%
  group_by(region,state) %>%
  summarize(total_pop=sum(population)) %>%
  summarize(average_pop=mean(total_pop), median_pop=median(total_pop))
```
Top n verb

```{r}
counties_selected <- counties %>%
  select(region,state,county,metro, population, walk)
counties_selected %>%
  group_by(region) %>%
  top_n(1, walk)
counties_selected <- counties %>%
  select(region,state,county, population, income)
counties_selected %>%
  group_by(region,state) %>%
  summarize(average_income = mean(income)) %>%
  top_n(1, average_income)
```

```{r}
counties_selected <- counties %>%
  select(state,metro, population)
counties_selected %>%
  group_by(state, metro) %>%
  summarize(total_pop=sum(population))%>%
  top_n(1,total_pop) %>%
  ungroup() %>%
  count(metro)
```

Selectin g columns

```{r}
counties %>%
  select(state,county,population, professional:production) %>%
  arrange(desc(service))
counties %>%
  select(state,county,population, ends_with("work")) %>%
  filter(public_work>=50)
counties %>%
  select(state,county,population, ends_with("at"))
```